On the computational complexity of detecting possibilistic locality

### 2017-06-19 paper.tex
Nicer definition of the locality decision problem.

The notion of signalling encapsulates the idea that a particular distribution could be used as a resource to superluminally signal information between the two parties. An example of a signalling situation is demonstrated in Figure \ref{sig}. We will focus on \emph{no-signalling} tables, which cannot be used for signalling; all quantum-mechanically accessible tables are of this type. In terms of how our tables are constructed, this means that if a sub-row contains a 1, then it must contain a 1 in every column; if a measurement outcome is possible, then it is possible no matter which measurement choice is made at the other site.
\begin{figure}
\begin{equation*}
\begin{array}{c| c c | c c |}
&&&&\\ \hline
&1 &1  &1 & 1\\
&0&0 &1 &1\\ \hline
\end{array} 
\end{equation*}
\caption{A simple signalling situation.}
\label{sig}
\end{figure}

@@

\begin{defn}[$(j,k)$-\textsc{PossLoc}]
The $k$-outcome Possibilistic locality decision problem, $(j,k)$-\textsc{PossLoc} is as follows: the problem instance is an element of $\{0,1\}^{j+k+n+m}$, where $j$ and $k$ are the maximum number of measurement outcomes for each measurement on the two subsystems being measured, and $n$ and $m$ are the number of measurements available at each location. This should be thought of as a data table where a 1 indicates a possible outcome and a 0 indicates an impossible outcome. In general, we will take $n=m$ to maintain a single scaling factor; in fact this change is without loss of generality. In the case where $j=k$, we will denote this problem merely $k$-\textsc{PossLoc}. While this is not essential, we will also require that the data tables obey a possibilistic no-signalling principle; this merely streamlines the wording of the theorems. 
\end{defn}
\begin{thm} \cite{Mans2011}
2-\textsc{PossLoc} is in \textbf{P}.
\end{thm}
\begin{proof}
This proof proceeds by showing that the only kind of possibilistic nonlocality that can occur in the two-outcome case is the one originally identified by Hardy \cite{Hard1993}; that is to say, a feature of the following kind, in which a blank space can be substituted either for a 0 or for a 1:
\begin{equation*}
\begin{array}{c| c c | c c |}
&&&&\\ \hline
&1 & & 0 & \\
&& & & \\ \hline
&0& & &  \\
&& & & 0 \\ \hline
\end{array}
\end{equation*}
We can quickly see that the highlighted one is impossible to extend to a ``deterministic grid'', \emph{i.e.} no single deterministic hidden variable can account for that possibility without also introducing an event that is impossible in the diagram. Since the appearance of this structure is equivalent to possibilistic nonlocality in this case, an algorithm to determine whether or not a given data table is possibilistically local or nonlocal reduces to checking each possible set of four contexts of this type, of which there are $n^2(n-1)^2/4$. Since checking for the appearance of such structures is possible in constant time, our algorithm runs in $O(n^4)$.
\end{proof}
However, this argument is specific to the case in which we have measurements with at most two outcomes on each side. As an illustration that there are quantumly-accessible nonlocal data tables that are not reducible to a fine-graining of a Hardy paradox, we will now demonstrate a novel possibilistic nonlocality scenario in the (2,3) case. This will be a generalisation of Hardy's proof of Bell's theorem.

>>>>

The notion of signalling encapsulates the idea that a particular distribution could be used as a resource to superluminally signal information between the two parties. An example of a signalling situation is demonstrated in Figure \ref{sig}. We will focus on \emph{no-signalling} tables, which cannot be used for signalling; all quantum-mechanically accessible tables are of this type. In terms of how our tables are constructed, this means that if a sub-row (sub-column) contains a 1, then it must contain a 1 in every column (row); if a measurement outcome is possible, then it is possible no matter which measurement choice is made at the other site.
\begin{figure}
\begin{equation*}
\begin{array}{c| c c | c c |}
&b_1=0&b_1=1&b_2=0&b_2=1\\ \hline
a_1=0&1 &1  &1 & 1\\
a_1=1&0&0 &1 &1\\ \hline
\end{array} 
\end{equation*}
\caption{A simple signalling situation; if Alice sees the outcome $a_1=1$, she knows Bob must have chosen measurement setting $B_2$.}
\label{sig}
\end{figure}

@@

\begin{defn}[$(l,m)$-\textsc{PossLoc}]
\quad\\
Instance: a possibilistic table of data represented by a  function $T:A\times B \times \{1,\dots, l\}\times \{1,\dots, m\}\rightarrow\{0,1\}$.

Question: do there exist a set of functions $T_i$, such that the $T_i$ are deterministic, and $T=\sum_i T_i$?
\end{defn}
Tihs represents a two-party nonlocality scenario in which Alice has access to measurements with $m$ outcomes, and Bob has access to measurements with $l$ outcomes. We are given a possibility table and asked whether or not it displays nonlocality. In general, we will take $|A|=|B|=n$ to maintain a single scaling factor; in fact this change is without loss of generality. In the case where $l=m$, we will denote this problem merely $l$-\textsc{PossLoc}. While this is not essential, we will also require that the data tables obey the possibilistic no-signalling principle; this merely streamlines the wording of the theorems. 
\begin{thm} \cite{Mans2011}
2-\textsc{PossLoc} is in \textbf{P}.
\end{thm}
\begin{proof}
This proof proceeds by showing that the only kind of possibilistic nonlocality that can occur in the two-outcome case is the one originally identified by Hardy \cite{Hard1993}; that is to say, a feature of the following kind, in which a blank space can be substituted either for a 0 or for a 1:
\begin{equation*}
\begin{array}{c| c c | c c |}
&&&&\\ \hline
&1 & & 0 & \\
&& & & \\ \hline
&0& & &  \\
&& & & 0 \\ \hline
\end{array}
\end{equation*}
We can quickly see that the highlighted one is impossible to extend to a ``deterministic grid'', \emph{i.e.} no single deterministic hidden variable can account for that possibility without also introducing an event that is impossible in the diagram. Since the appearance of this structure is equivalent to possibilistic nonlocality in this case, an algorithm to determine whether or not a given data table is possibilistically local or nonlocal reduces to checking each possible set of four contexts of this type, of which there are $n^2(n-1)^2/4$. Since checking for the appearance of such structures is possible in constant time, our algorithm runs in $O(n^4)$.
\end{proof}
However, this argument is specific to the case in which we have measurements with at most two outcomes on each side. As an illustration that there are quantumly-accessible nonlocal data tables that are not reducible to a fine-graining of a Hardy paradox, we will now demonstrate a novel possibilistic nonlocality scenario in the (2,3) case. This will be a generalisation of Hardy's proof of Bell's theorem.
